---
title: "Data Wrangling"
author: "Ajay Koli"
output:
  html_document:
    theme: united
    highlight: tango
    toc: TRUE
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: TRUE
    #code_folding
---
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE, comment = NA, warning = FALSE, errors = FALSE, message = FALSE, tidy = FALSE, cache = FALSE, fig.path= "03-figs/")
```

# data wrangling

it combines data tidying and transformation.

## data tidying

## data transformation


# Packages required

```{r}
# to use the dplyr pkg
library(tidyverse)
```

## tidyverse 

It is a set of these packages:

- `ggplot2`, for data visualisation.

- `dplyr`, for data manipulation.

- `tidyr`, for data tidying.

- `readr`, for data import.

- `purrr`, for functional programming.

- `tibble`, for tibbles, a modern re-imagining of data frames.

- `stringr`, for strings.

- `forcats`, for factors.

We will be using dplyr and readr packages.

## palmerpenguins

```{r}
# to get penguins data
library(palmerpenguins)
```

and from this package we will be using `penguins` data

```{r}
penguins
```

### types of the variables

- `int` stands for integers.

- `dbl` stands for doubles, or real numbers.

- `chr` stands for character vectors, or strings.

- `dttm` stands for date-times (a date + a time).

- `lgl` stands for logical, vectors that contain only TRUE or FALSE.

- `fct` stands for factors, which R uses to represent categorical variables with fixed possible values.

- `date` stands for dates.


as we can see in the output all variable names are not visible. we can use a specific R function to just read the name of all the variables in the data set.

```{r}
names(penguins)
```

we can use the help function to learn more about the penguins data

```{r}
?penguins
```

# dplyr package

five key functions:

- Pick observations by their values using function `filter()`.
    
- Reorder the rows using function `arrange()`.

- Pick variables by their names using function `select()`.

- Create new variables with functions of existing variables using function `mutate()`.

- Collapse many values down to a single summary using function `summarise()`.

## 
